---
title: "DiffBind TbsP"
author: "Rylee Hackley"
project: "HVO trmB and tbsP"
output: github_document
---
```{r, message=FALSE}
# if using windows and getting an rJava error when intalling xlsx package, need to make sure that your java matches r installations (32 or 64-bit) #Sys.setenv(JAVA_HOME="C:\\Program Files/Java/jre1.8.0_241/")
# library(rJava)
# BiocManager::install(c("DiffBind", "edgeR"))
library(tidyverse)
library(openxlsx)
library(rtracklayer)
library(DiffBind)

package.version("DiffBind")
```

fragment sizes calculated by CHIPQC. using initial peaks called with default settings, since PCA had better replicate agreement and %RIP was higher
IP_tbsPHA_1	213
IP_tbsPHA_2	236
IP_tbsPHA_3	214
IP_tbsPHA_4	233
IP_tbsPHA_glc_1	243 (very low RIP, might remove as outlier)
IP_tbsPHA_glc_2	231
IP_tbsPHA_glc_3	221
IP_tbsPHA_glc_4	223
IP_WT_glc_2	201

```{r, message=F}
# load all samples
meta <- read_csv("01b_ChIPQC/HVO_chipqc_meta_initial.csv")[c(2:9),]
config <- list(minQCth = 30, fragmentSize = c(243, 231, 221, 223, 213, 236, 214, 233), bCorPlot = T, th = 0.05)

grey <- import.bed("01a_mosaics/peaks/HVO_WT_glc_2.bed")
```

Create initial DBA object
```{r, message=F, fig.height=6, fig.width=6}
tbsP <- dba(sampleSheet = meta, peakFormat = "bed", minOverlap = 0.75, config = config)
tbsP
```

```{r, fig.height=6, fig.width=6}
plot(tbsP)

# number of overlapping peaks in samples
dba.overlap(tbsP, mode = DBA_OLAP_RATE)
dba.overlap(tbsP, tbsP$masks$`0.1% glucose`, mode = DBA_OLAP_RATE)
dba.overlap(tbsP, tbsP$masks$`no glucose`, mode = DBA_OLAP_RATE)
```

Mask peaks identified in WT control samples
```{r, fig.height=6, fig.width=6}
tbsP <- dba.blacklist(tbsP, blacklist = FALSE, greylist = grey)
dba.overlap(tbsP, mode = DBA_OLAP_RATE)
dba.overlap(tbsP, tbsP$masks$`0.1% glucose`, mode = DBA_OLAP_RATE)
dba.overlap(tbsP, tbsP$masks$`no glucose`, mode = DBA_OLAP_RATE)
```

```{r}
tbsP <- dba.count(tbsP, summits = 150, minOverlap = 0.75) # DBA_SCORE_SUMMIT_ADJ: summit height (maximum read pileup value), normalized to relative library size
tbsP
```

compare normalization:
```{r}
tbsP2 <- dba.normalize(tbsP, normalize = DBA_NORM_LIB)
tbsP2 <- dba.contrast(tbsP2, reorderMeta = list(Condition = "0.1% glucose"), minMembers = 3)
tbsP2 <- dba.analyze(tbsP2)
dba.plotMA(tbsP2, method = DBA_DESEQ2)

tbsP4 <- dba.normalize(tbsP, normalize = DBA_NORM_LIB, library = DBA_LIBSIZE_PEAKREADS)
tbsP4 <- dba.contrast(tbsP4, reorderMeta = list(Condition = "0.1% glucose"), minMembers = 3)
tbsP4 <- dba.analyze(tbsP4)
dba.plotMA(tbsP4, method = DBA_DESEQ2)

tbsP3 <- dba.normalize(tbsP, normalize = DBA_NORM_RLE, library = DBA_LIBSIZE_PEAKREADS)
tbsP3 <- dba.contrast(tbsP3, reorderMeta = list(Condition = "0.1% glucose"), minMembers = 3)
tbsP3 <- dba.analyze(tbsP3)
dba.plotMA(tbsP3, method = DBA_DESEQ2)
```
DBA norm RLE is the only normalization that results in correct clustering: 12 DB peaks.

normalize, set up contrasts, and execute differential binding analysis
```{r}
tbsP <- dba.normalize(tbsP, normalize = DBA_NORM_LIB, library = DBA_LIBSIZE_FULL)
tbsP <- dba.contrast(tbsP, reorderMeta = list(Condition = "0.1% glucose"), minMembers = 3)
tbsP <- dba.analyze(tbsP)

dba.plotMA(tbsP, method = DBA_DESEQ2)
dba.show(tbsP, bContrasts = TRUE)
```

Visualize DBA results
```{r}
dba.plotHeatmap(tbsP3, contrast = 1)

# PCA for all sites in affinity matrix
dba.plotPCA(tbsP3, DBA_CONDITION, label = DBA_CONDITION)

# PCA plot using affinity data for only differential bound sites
dba.plotPCA(tbsP3, contrast = 1, label = DBA_CONDITION)

# Binding affinity heatmap showing affinities for differentially bound sites
hmap <- colorRampPalette(c("red", "black", "green"))(n = 13)
dba.plotHeatmap(tbsP3,
  contrast = 1,
  correlations = FALSE, scale = "row", colScheme = hmap
)
```

view report, compile score, and export
```{r}
(tbsP.DB <- dba.report(tbsP3, bFlip = T, bCalled = T, bCounts = T, bNormalized = F))
tbsP.DB$score <- tbsP.DB$Fold

export.bed(tbsP.DB, "02_DiffBind/02_diffbind_RLE.bed")
write_csv(as.data.frame(tbsP.DB), "02_DiffBind/02_diffbind_RLE.csv")
```
