---
title: "Binding TbsP"
author: "Rylee Hackley"
project: "HVO trmB and tbsP"
output: github_document
---
```{r, message=FALSE}
# if using windows and getting an rJava error when intalling xlsx package, need to make sure that your java matchs r installations (32 or 64-bit) #Sys.setenv(JAVA_HOME="C:\\Program Files/Java/jre1.8.0_241/")
# library(rJava)

library(tidyverse)
library(openxlsx)
library(rtracklayer)
library(DiffBind)

package.version("DiffBind")
```

fragment sizes calculated by CHIPQC
IP_tbsPHA_1	213
IP_tbsPHA_2	236
IP_tbsPHA_3	214
IP_tbsPHA_4	233
IP_tbsPHA_glc_1	243
IP_tbsPHA_glc_2	231
IP_tbsPHA_glc_3	221
IP_tbsPHA_glc_4	223
IP_WT_glc_2	201

```{r, message=F}
(meta <- read_csv("01b_ChIPQC/HVO_chipqc_meta_initial.csv")[c(2:9), ])

config <- list(minQCth = 30, fragmentSize = c(243, 231, 221, 223, 213, 236, 214, 233), bCorPlot = T, th = 0.05)

grey <- import.bed("01a_mosaics/peaks/HVO_WT_glc_2.bed")
```

Create initial DBA object
```{r, message=F, warning=FALSE, fig.height=6, fig.width=6}
tbsP <- dba(sampleSheet = meta, peakFormat = "bed", minOverlap = 0.75, config = config)

# Mask peaks identified in WT control samples
tbsP2 <- dba.blacklist(tbsP, blacklist = FALSE, greylist = grey)
```

```{r, fig.height=6, fig.width=6}
plot(dba.overlap(tbsP, mode = DBA_OLAP_RATE), type = "b", ylab = "# peaks", xlab = "Overlap at least this many peaksets")

# number of overlapping peaks in samples
dba.overlap(tbsP, mode = DBA_OLAP_RATE)
dba.plotVenn(tbsP, tbsP$masks$`no glucose`)
dba.plotVenn(tbsP, tbsP$masks$`0.1% glucose`)

# consensus peaksets
tbsP_consensus <- dba.peakset(tbsP, consensus = DBA_CONDITION, minOverlap = 0.75)

dba.plotPCA(tbsP_consensus, DBA_CONDITION, label = DBA_ID)


dba.plotVenn(tbsP_consensus, tbsP_consensus$masks$Consensus) # 32 regions in no_glu (onlyA)

(tbsP_consensus.OL <- dba.overlap(tbsP_consensus, tbsP_consensus$masks$Consensus))

# get the combined gRanges:
dba.peakset(tbsP_consensus, bRetrieve = T)
```
This doesn't have any peak information since it is solely looking at overlaps. In order to trim and recalculate consensus peaks, need to do a hybrid approach.

##tbsP no_glu
```{r}
meta %>% filter(Condition == "no glucose") -> meta.noglu
config <- list(
  minQCth = 30, fragmentSize = c(213, 236, 214, 233),
  bCorPlot = T, th = 0.05
)
tbsP.noglu <- dba(sampleSheet = meta.noglu, peakFormat = "bed", minOverlap = 0.75, config = config)
tbsP.noglu <- dba.blacklist(tbsP.noglu, blacklist = FALSE, greylist = grey)
count.noglu <- dba.count(tbsP.noglu, summits = 150, minOverlap = 0.75, score = DBA_SCORE_RPKM_MINUS) # 42 peaks across 4 samples
count.noglu <- dba.normalize(count.noglu, normalize = DBA_NORM_RLE, library = DBA_LIBSIZE_FULL)
as.data.frame(peak.noglu <- dba.peakset(count.noglu, bRetrieve = T))
dba.plotVenn(tbsP.noglu, tbsP.noglu$masks$`no glucose`)
```

##tbsP glu
```{r}
meta %>% filter(Condition == "0.1% glucose") -> meta.glu
config <- list(
  minQCth = 30, fragmentSize = c(243, 231, 221, 223),
  bCorPlot = T, th = 0.05
)
tbsP.glu <- dba(sampleSheet = meta.glu, peakFormat = "bed", minOverlap = 0.75, config = config)
tbsP.glu <- dba.blacklist(tbsP.glu, blacklist = FALSE, greylist = grey)
count.glu <- dba.count(tbsP.glu, summits = 150, minOverlap = 0.75, score = DBA_SCORE_RPKM_MINUS) # returned 45 peaks across 4 samples
count.glu <- dba.normalize(count.glu, normalize = DBA_NORM_RLE, library = DBA_LIBSIZE_FULL)
count.glu$peaks
as.data.frame(peak.glu <- dba.peakset(count.glu, bRetrieve = T))
dba.plotVenn(tbsP.glu, tbsP.glu$masks$`0.1% glucose`)
```

##shared peaks calculated by ignoring condition differences:
```{r}
tbsP <- dba(sampleSheet = meta, peakFormat = "bed", minOverlap = 0.75, config = config)
tbsP <- dba.blacklist(tbsP, blacklist = FALSE, greylist = grey)
counted <- dba.count(tbsP, summits = 150, minOverlap = 0.75, score = DBA_SCORE_RPKM_MINUS) # returned 45 peaks across 4 samples
count.shared <- dba.normalize(counted, normalize = DBA_NORM_RLE, library = DBA_LIBSIZE_FULL)
count.shared$peaks
as.data.frame(shared <- dba.peakset(count.shared, bRetrieve = T))
shared$score <- rowMeans(as.data.frame(mcols(shared)))

dba.plotVenn(tbsP, tbsP$masks$Replicate.1)
dba.plotVenn(tbsP, tbsP$masks$Replicate.2)
dba.plotVenn(tbsP, tbsP$masks$Replicate.3)
dba.plotVenn(tbsP, tbsP$masks$Replicate.4)
```

Now, remove peaks in "unique" lists that overlap with the shared list.
```{r}
# unique no_glu
peak.noglu.unique <- peak.noglu[-queryHits(findOverlaps(peak.noglu, shared, type = "any", minoverlap = 100)), ]
peak.noglu.unique$score <- rowMeans(as.data.frame(mcols(peak.noglu.unique)))
peak.noglu.unique <- sort(peak.noglu.unique, by = ~score, decreasing = T)
names(peak.noglu.unique) <- seq(length(peak.noglu.unique))
peak.noglu.unique <- sort(peak.noglu.unique)
as.data.frame(peak.noglu.unique)

# unique glu
peak.glu.unique <- peak.glu[-queryHits(findOverlaps(peak.glu, shared, type = "any", minoverlap = 100)), ]
peak.glu.unique$score <- rowMeans(as.data.frame(mcols(peak.glu.unique)))
peak.nlu.unique <- sort(peak.glu.unique, by = ~score, decreasing = T)
names(peak.glu.unique) <- seq(length(peak.glu.unique))
peak.glu.unique <- sort(peak.glu.unique)
as.data.frame(peak.glu.unique)

# shared
shared <- sort(shared, by = ~score, decreasing = T)
names(shared) <- seq(length(shared))
shared <- sort(shared)
as.data.frame(shared)

# prepare for exporting (bed)
export.bed(peak.glu.unique, "02_DiffBind/glucose_peaks.bed") # includes score
export.bed(peak.noglu.unique, "02_DiffBind/noglucose_peaks.bed") # includes score
export.bed(shared, "02_DiffBind/tbsP_shared_peaks.bed") # includes score
```

Compare with affinity analysis
```{r}
DB <- import.bed("02_DiffBind/02_diffbind_RLE.bed")
DB <- DB[, 2]

peak.noglu.unique[queryHits(findOverlaps(peak.noglu.unique, DB, type = "any", minoverlap = 100)), ]
peak.glu.unique[queryHits(findOverlaps(peak.glu.unique, DB, type = "any", minoverlap = 100)), ]

shared[queryHits(findOverlaps(shared, DB, type = "any", minoverlap = 100)), ] # all 11 of DB calls are in the shared dataset
DB[queryHits(findOverlaps(DB, shared, type = "any", minoverlap = 100)), ]
```
all 11 of DB peaks are in the shared peak set. Since none of the peaks are greatly differentially bound (max 0.4 LFC), I am not concerned with wrangling the data to include the LFC for those peaks in the same sheet.

Add analysis type metadata
```{r}
peak.noglu.unique$analysis <- "occupancy"
peak.glu.unique$analysis <- "occupancy"
shared$analysis <- "occupancy"
grey$analysis <- "GreyListChIP"
DB$analysis <- "affinity"
```

Compile Supplemental file:
```{r}
variables <- GRangesList(
  "tbsP_shared" = shared, "tbsP_DiffBind" = DB,
  "no_glu_unique" = peak.noglu.unique, "glu_unique" = peak.glu.unique,
  "WT_greylist" = grey
)
wb <- createWorkbook()

for (i in 1:length(variables)) {
  # create dataframe
  df <- as.data.frame(variables[i]) %>% discard(~ all(is.na(.)))
  df <- df[-1]

  # write to excel file, storing results for each .bed as a different sheet
  addWorksheet(wb, names(variables)[i], gridLines = TRUE) # Add a worksheet
  writeData(wb, sheet = i, df, rowNames = FALSE) # write data to worksheet number i
}

saveWorkbook(wb, "02_DiffBind/02_peaklists.xlsx", overwrite = TRUE)
```
