---
title: "2022-05-02-bioscreen-tbsP-AKS"
author: "Angie Vreugdenhil-Hayslette & Amy Schmid"
date: '`r Sys.Date()`'
output: html_document
---

```{r}
# fresh start
rm(list = ls()) 
```

```{r}
# load required libraries
library(plyr)
library(tidyr) 
library(dplyr)
library(tibble)
library(viridis)
library(ggplot2)
library(tinytex)
library(latexpdf)
library(data.table)
library(growthrates)
library(rstatix)
library(ggpubr)
library(EnvStats)
```

## SETTINGS
```{r}
##################################################################################################### SETTINGS
# project folder
folder <- '~/Dropbox (Duke Bio_Ea)/projects/TbsP/data/SH-suppressor-data/'
# folder with meta
foldermeta <- '~/Dropbox (Duke Bio_Ea)/projects/TbsP/data/SH-suppressor-data/meta/'
# metafile name
meta <- '2022-05-02-tbsP-all-meta.csv'
# folder with data
folderdata <- '~/Dropbox (Duke Bio_Ea)/projects/TbsP/data/SH-suppressor-data/data/'
# datafile name
data <- '2022-05-03-tbsP-all-data.csv'
# wells to be removed based on observation (do not remove = 1000)
#remove1<- 79
#remove2<- 170
#remove3<- 197
#remove4<- 1000
#remove5<- 1000
#remove6<- 1000
#remove7<- 1000
#remove8<- 1000
#remove9<- 1000
#remove10<- 1000
#remove11<- 1000
#remove12<- 1000
# blank wells (not a blank = 1000)
#blank1<- 20
#blank2<- 52
#blank3<- 90

```

## read in RAW DATA
```{r}

##################################################################################################### RAW DATA
# load data
## read in metafile
mt <- read.csv(paste(foldermeta, meta, sep = ''),
               sep = ',',
               header = TRUE)
mt0 <- mt[-1, -(1:2)]
## read in datafile
dt <- read.csv(paste(folderdata, data, sep = ''),
               sep = ',',
               header = FALSE)
dt0 <- dt[-(1:2), -2 ]
colnames(dt0) <- c('time', dt0[1, -1])
dt0 <- dt0[-1, ]

# remove rows based on observation in meta
#mt0 <- mt0[!(mt0$variable == remove1 |
 #              mt0$variable == remove2 |
  #             mt0$variable == remove3 |
   #            mt0$variable == remove4 |
    #           mt0$variable == remove5 |
     #          mt0$variable == remove6 |
      #         mt0$variable == remove7 |
       #        mt0$variable == remove8 |
        #       mt0$variable == remove9 |
         #      mt0$variable == remove10 |
          #     mt0$variable == remove11 |
            #   mt0$variable == remove12),]

# remove rows based on observation in data
#dt0 <- dt0 %>% select(-contains(as.character(remove1)))
#dt0 <- dt0 %>% select(-contains(as.character(remove2)))
#dt0 <- dt0 %>% select(-contains(as.character(remove3)))
```

## Format data
```{r}
# extract hours, mins, secs (Comment in if your time is listed in format of "1:30:06". Not needed if time is already expressed in hours (e.g. "1.5" = 1 hr 30 mins))
#hou <-
#  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 1))
#min <-
#  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 2))
#sec <-
#  as.numeric(lapply(strsplit(as.character(dt0$time), ':'), '[[', 3))
## set total time to hours
#dt0$time <- (hou) + min / 60 + (sec / 3600)

# melt the dataframe to one list
m_dt <- melt(dt0, id = 'time')
## merge
mtdt <- merge(m_dt, mt0, by = 'variable')
mtdt$environment <- ifelse(mtdt$condition=='',as.character(mtdt$media),paste(mtdt$media,'+',mtdt$condition,sep=''))
mtdt$ID <- ifelse(mtdt$condition=='',as.character(mtdt$name),paste(mtdt$name,' (',mtdt$condition,')',sep=''))
well <- paste(mtdt$variable, mtdt$strain, sep = ' ')
mtdt <- cbind(well, mtdt)
## remove rows without ID
c_mtdt <- mtdt[!(mtdt$ID == ''), ]
```

#Time cutoff (comment out if not needed)
```{r}
# time cutoff
t_mtdt <- c_mtdt#[!(c_mtdt$time <= 0.5 | c_mtdt$time >= 48), ]
t_mtdt[500:510,] #look at the data to make sure time cutoff worked properly
## fix delta symbol
#t_mtdt$ID<- gsub('_','\u0394',t_mtdt$ID)
#t_mtdt$name<- gsub('_','\u0394',t_mtdt$name)
## select blanks (three tech reps each)
#blanks<- t_mtdt[(t_mtdt$variable == 1 |
 #                  t_mtdt$variable == 2 |
  #               t_mtdt$variable == 3 |
   #              t_mtdt$variable == 4 |
    #               t_mtdt$variable ==  5 |
     #              t_mtdt$variable == 6 |
      #             t_mtdt$variable == 7 |
       #            t_mtdt$variable == 8 |
        #           t_mtdt$variable == 9 |
         #          t_mtdt$variable == 10 |
          #         t_mtdt$variable == 11 |
           #        t_mtdt$variable == 12), ]
```

## Calculate stats
```{r}
## summarize all values (time, strain, media, condition, biorep)
tm_mtdt<- ddply(t_mtdt, c('time', 'ID', 'strain', 'name', 'media', 'condition', 'environment', 'biorep'),
               summarize,
               value_mn = mean(as.numeric(value)),
               value_sd = sd(value))
```

## If you included blank wells, subtract blanks from culture OD readings
```{r}
# set baseline (blank wells)
## take minimal value of each blank per condition
#blank_all<- ddply(blanks, c('environment','time'),
       #       summarize,
        #     blank=min(value)-0.01)
##Comment "blank_all" in if you used blanks.

## add correct blank values to each row
#bu_mtdt<- merge(t_mtdt,blank_all,c('environment','time')) # use this if you have blank wells
bu_mtdt<- merge(t_mtdt,c('environment','time')) #use this if you don't have blank wells
## reorganize columns
#bu_mtdt$value_mn<- bu_mtdt$value - bu_mtdt$blank  #subtract blank values if you have blanks
bu_mtdt$value_mn<- bu_mtdt$value #use this if you don't have blanks
#bu_mtdt$blank<- NULL
## summarize all blank corrected values (time, strain, media, condition, biorep)
bm_mtdt<- ddply(bu_mtdt, c('time', 'ID', 'strain', 'name', 'media', 'condition', 'environment', 'biorep'),
                summarize,
                value_mn = mean(value),
                value_sd = sd(value))
## remove blank rows
b_mtdt <- bu_mtdt[!bu_mtdt$name=='blank',]
bm_mtdt <- bm_mtdt[!bm_mtdt$name=='blank',]
min(bm_mtdt$value_mn)
min(bm_mtdt$value_mn)

```

## PLOT BLANKS - only needed if you included blank wells
```{r}
##################################################################################################### PLOT BLANKS
# all growth curves
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))
# blanks
p_blanks<- 
  ggplot(blanks, aes(x = time, y = value, color = environment)) +
  ggtitle('Blank curves') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('condition', discrete = TRUE) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=10, angle=0))

ggsave(p_blanks,
       file = paste(
         format(Sys.time(), '%Y%m%d'),
         '_',
         'plot_blanks',
         #'.jpeg',
         '.pdf',
         sep = ''),
       width = 10, height = 8, units = "in", device='pdf')

```

## PLOT RAW
```{r}
##################################################################################################### PLOT RAW
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# growthcurves - one plot - well
#p_well <- 
#pdf('tbsP_rawplots.pdf', width = 6, height = 8)
  ggplot(t_mtdt, aes(x = as.numeric(time), y = as.numeric(value))) +
  ggtitle('Growthcurves each well') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('well', discrete = TRUE) +
  theme(legend.position = 'none',
        strip.text = element_text(face = "italic",size = 8),
        plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
        axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
        axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
        legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
        axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
        axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
        legend.text = element_text(face='plain', color='#000000', size=10, angle=0))
#dev.off()
#ggsave(p_well, file = paste(
 # format(Sys.time(), '%Y%m%d'),
  #'_',
  #'plot_well_all',
 # '.jpeg',
 #'.pdf',
#  sep = ''),
 # width = 10, height = 8, units = "in", device='pdf')
```

```{r}
# growthcurves - one plot - ID
p_id <- 
  ggplot(t_mtdt, aes(x = as.numeric(time), y = as.numeric(value), color = ID)) +
  ggtitle('Growthcurves each strain') +
  xlab('time (h)') + 
  ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_point() +
  scale_y_continuous(trans = 'log2') +
  scale_color_viridis('strain description', discrete = TRUE) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_id, file = paste(
  format(Sys.time(), '%Y%m%d'),
  '_',
  'plot_ID_all',
  #'.jpeg',
  '.pdf',
  sep = ''),
  width = 10, height = 8, units = "in", device='pdf')
```

```{r}
# growthcurves - grid
p_gri <-
  ggplot(t_mtdt, aes(x = as.numeric(time), y = as.numeric(value))) +
  ggtitle('Growthcurves each well') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_wrap(~ well) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "plain",size = 8),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=8, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=8, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=8, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=6, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=6, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=8, angle=0))

ggsave(p_gri, file = paste(
  format(Sys.time(), '%Y%m%d'),
  '_',
  'plot_grid_all',
  '.jpeg',
  sep = ''),
  width = 10, height = 8, units = "in", device='jpeg')
```

```{r}

##################################################################################################### PLOT STATS
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

# 95% confidence interval function
conf_int95 <- function(data) {
  n <- length(data)
  error <- qt(0.975, df = n - 1) * sd(data) / sqrt(n)
  return(error)
}

stats <- tm_mtdt %>%
  group_by(ID, strain, name, environment, time) %>%
  summarise(
    reps = length(value_mn),
    average = mean(value_mn),
    CI95 = conf_int95(value_mn)) %>%
  filter(!strain == '')


# growth curves - average strain
p_ave_str <-
ggplot(stats, aes(x = as.numeric(time), y = average, color = environment)) +
  ggtitle('Growthcurves averaged each condition and separated by strain') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI95,
      ymax = average + CI95,
      fill = environment), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(name ~ .) +
  scale_fill_viridis('condition', discrete = TRUE) +
  scale_color_viridis('condition', discrete = TRUE) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 7),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='plain', color='#000000', size=10, angle=0))

ggsave(p_ave_str,
       file = paste(
         format(Sys.time(), '%Y%m%d'),
         '_',
         'plot_average_strain',
         '.pdf',
         sep = ''),
       width = 4, height =6, units = "in", device='pdf')


```

```{r}
# growth curves - average condition
p_ave_con <-
ggplot(stats, aes(x = as.numeric(time), y = average, color = name)) +
  ggtitle('Growthcurves averaged each strain and separated by condition') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI95,
      ymax = average + CI95,
      fill = name), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(environment ~ .) +
  scale_fill_viridis('Strain name', discrete = TRUE) +
  scale_color_viridis('Strain name', discrete = TRUE) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 10),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_ave_con,
      file = paste(
        format(Sys.time(), '%Y%m%d'),
        '_',
        'plot_average_condition',
        '.pdf',
        sep = ''),
      width = 4, height = 6, units = "in", device='pdf')
```

## plot average conditions, 0.05% glucose only - Fig 5A
```{r}
#set the colors
group.colors <- c(pyrE = '#440154', trmB="#35B779", trmBtbsP ="#A6A560", tbsP ="#E79A51")

#filter the data to include only 0.05% glucose
stats.05 <- filter (stats, ! ID == 'trmB (glu01)', ! ID == 'pyrE (glu01)', ! ID == 'tbsP (glu01)', ! ID == 'trmBtbsP (glu01)')

#plot the data and export to pdf file
pdf('growthcurve-stats.05.pdf', height = 5, width = 5)
ggplot(stats.05, aes(x = as.numeric(time), y = average, color = name)) +
  ggtitle('Growthcurves averaged each strain and separated by condition') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI95,
      ymax = average + CI95,
      fill = name), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(environment ~ .) +
  scale_fill_manual(values = group.colors) +
  scale_color_manual(values = group.colors) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 10),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))
dev.off()
```


```{r}
##################################################################################################### MUMAX Growthrates
# folder for plots
dir.create(paste(folder,'plots',sep=''))
setwd(paste(folder,'plots',sep=''))

t_mtdt$time <- as.numeric(t_mtdt$time)
t_mtdt$value <- as.numeric(t_mtdt$value)

# extract mumax for plotting fit
pretty_spline_fits <-
  all_splines(value ~ time | ID + biorep,
              data = t_mtdt,
              spar = 0.5)
# extract mumax for table
all_spline_fits <-
  all_splines(value ~ time | ID + biorep + strain + name + environment,
              data = t_mtdt,
              spar = 0.5)

#plot spline fit
jpeg(paste(
  format(Sys.time(), '%Y%m%d'),
  '_',
  'plot_splines',
  '.jpeg',sep=''),
  width = 1188, 
  height = 880)
par(mfrow = c(8, 7))
par(mar = c(2, 2, 1.5, 1.25))
plot(pretty_spline_fits, log='y')
dev.off()

# store results in dataframe
mumax<- results(all_spline_fits)
row.names(mumax)<- NULL

# remove blanks
mumax <- mumax[!mumax$strain=='blank',]


# remove mumax with rsquared below 0.95
mumax$mumax <- ifelse(mumax$r2 < 0.95, 0, mumax$mumax)
#mumax.noHA <-  ifelse(mumax.noHA$r2 < 0.95, 0, mumax.noHA$mumax)
# calculate mean mumax
mumean<- ddply(mumax,c('ID','strain','name','environment'), 
               summarize,
               OD_mn=mean(y0),
               OD_sd=sd(y0),
               mumax_mn=mean(mumax),
               mumax_sd=sd(mumax))

```



```{r}
#reorder
mumean<- mumean[order(mumean$mumax_mn),]
mumean$ID <- factor(mumean$ID, levels = mumean$ID[order(mumean$mumax_mn)])

#mumean.noHA<- mumean.noHA[order(mumean.noHA$mumax_mn),]
#mumean.noHA$ID <- factor(mumean.noHA$ID, levels = mumean.noHA$ID[order(mumean.noHA$mumax_mn)])
```

```{r}
# growthcurves - bars
p_mu <- 
  ggplot(data = mumean, aes(x = ID, y = mumax_mn, fill = name)) +
  ggtitle('Mean mumax of bioreps comparison') +
  xlab('Strain name (condition)') + 
  ylab('[mumax]')+
  geom_bar(stat='identity') +
  geom_errorbar(aes(ymin=mumax_mn-mumax_sd, ymax=mumax_mn+mumax_sd), width=.2,
              position=position_dodge(.9))+
  coord_flip() +
   scale_fill_viridis('name', discrete = TRUE) +
  scale_color_viridis('name', discrete = TRUE) +
  theme(#legend.position = 'none',
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='italic', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))

ggsave(p_mu,
       file = paste(
         format(Sys.time(), '%Y%m%d'),
         '_',
         'plot_mumax',
         '.jpeg',
         sep = ''),
       width = 10, height = 8, units = "in", device='jpeg')
```

## Plot mumax data, excluding 0.01% condition (does not fully rescue phenotype - 0.05% rescues)
```{r}
mumean.05<- filter (mumean, ! ID == 'trmB (glu01)', ! ID == 'pyrE (glu01)', ! ID == 'tbsP (glu01)', ! ID == 'trmBtbsP (glu01)')



pdf('mumean.05.pdf', height = 4, width = 6)
  ggplot(data = mumean.05, aes(x = ID, y = mumax_mn, fill = name)) +
  ggtitle('Mean mumax of bioreps comparison') +
  xlab('Strain name (condition)') + 
  ylab('[mumax]')+
  geom_bar(stat='identity') +
  geom_errorbar(aes(ymin=mumax_mn-mumax_sd, ymax=mumax_mn+mumax_sd), width=.2,
              position=position_dodge(.9))+
  coord_flip() +
   scale_fill_manual(values = group.colors) +
  scale_color_manual(values = group.colors) +
  theme(#legend.position = 'none',
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='italic', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))
dev.off()
#ggsave(p_mu.05only,
 #      file = paste(
  #       format(Sys.time(), '%Y%m%d'),
   #      '_',
    #     'plot_mumax_05',
     #    '.pdf',
      #   sep = ''),
      # width = 10, height = 8, units = "in", device='jpeg')

```

## inset, fig 5A, tbsP vs parent
```{r}
stats.05 <- filter (stats, ! ID == 'trmB (glu01)', ! ID == 'pyrE (glu01)', ! ID == 'tbsP (glu01)', ! ID == 'trmBtbsP (glu01)') %>%
  filter (name == "pyrE" | name == "trmBtbsP")

#plot the data and export to pdf file
pdf('growthcurve-stats.05.inset.pdf', height = 5, width = 5)
ggplot(stats.05, aes(x = as.numeric(time), y = average, color = name)) +
  ggtitle('Growthcurves averaged each strain and separated by condition') +
  xlab('time (h)') + 
 ylab(bquote(bold(paste('Log'['2'],'(Absorbance at 600nm)')))) +
  geom_ribbon(
    aes(
      ymin = average - CI95,
      ymax = average + CI95,
      fill = name), color = NA, alpha = 0.3) +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  facet_grid(environment ~ .) +
  scale_fill_manual(values = group.colors) +
  scale_color_manual(values = group.colors) +
  theme(#legend.position = 'none',
    strip.text = element_text(face = "italic",size = 10),
    plot.title = element_text(face='bold', color='#000000', size=12, angle=0),
    axis.title.x = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.title.y = element_text(face='bold', color='#000000', size=10, angle=90),
    legend.title = element_text(face='bold', color='#000000', size=10, angle=0),
    axis.text.y = element_text(face='plain', color='#000000', size=8, angle=0),
    axis.text.x = element_text(face='plain', color='#000000', size=8, angle=0),
    legend.text = element_text(face='italic', color='#000000', size=10, angle=0))
dev.off()
```

## is the growth of ∆tbsP and ∆trmB∆tbsP deletion strains significantly different from parent strain pyrE?


```{r}
	#Hypotheses to test (these are each independent hypotheses. don't need ANOVA, don't need multiple hypothesis testing, also cuz not so many tests):
	#1. Null hypothesis 1 = pyrE growth is not different from trmB in - glu
	#2. Null hypothesis 2 = pyrE growth is not different from tbsP in - glu
	#3. Null hypothesis 3 =  pyrE growth is not different from double mutant in -glu.
  #4. Null hypothesis 4 = pyrE growth is not different from tbsP in + glu
  #5. Null hypothesis 5 = pyrE growth in - glu is not different from pyrE in +glu
  #6. Null hypothesis 6 = trmB growth in - glu is not different from trmB in +glu
  #7. Null hypothesis 7 = tbsP growth in - glu is not different from tbsP in +glu
  #8. Null hypothesis 8 = trmBtbsP growth in - glu is not different from trmBtbsP in +glu
  #9. Null hypothesis 9 = pyrE growth in +glu is not different from trmB in +glu

hyp1 = filter (mumax, environment == "HvCA_ura+noGlu") %>%
  filter (name == "pyrE" | name == "trmB") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")


hyp2 = filter (mumax, environment == "HvCA_ura+noGlu") %>%
  filter (name == "pyrE" | name == "tbsP") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")


hyp3 = filter (mumax, environment == "HvCA_ura+noGlu") %>%
  filter (name == "pyrE" | name == "trmBtbsP") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")



hyp4 = filter (mumax, environment == "HvCA_ura+glu05") %>%
  filter (name == "pyrE" | name == "tbsP") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")



hyp5 = filter (mumax, ! environment   == "HvCA_ura+glu01") %>%
  filter (name == "pyrE") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")

hyp6 = filter (mumax, ! environment   == "HvCA_ura+glu01") %>%
  filter (name == "trmB") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")

hyp7 = filter (mumax, ! environment   == "HvCA_ura+glu01") %>%
  filter (name == "tbsP") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")

hyp8 = filter (mumax, ! environment   == "HvCA_ura+glu01") %>%
  filter (name == "trmBtbsP") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")

hyp9 = filter (mumax, environment == "HvCA_ura+glu05") %>%
  filter (name == "pyrE" | name == "trmB") %>%
  t_test(mumax ~ ID, p.adjust.method = "BH")

all_tests <- rbind(hyp1, hyp2, hyp3, hyp4, hyp5, hyp6, hyp7, hyp8, hyp9)
write.table(all_tests, file = "t-test_hypotheses_1to9.txt", sep = "\t")

#reject null for all hypotheses except #9.
```





##################################################################################################### END

