---
title: "R Notebook"
author: "Rylee Hackley & Amy K. Schmid"
project: "HVO trmB and tbsP"
output: github_document
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggpattern)
library(ggpubr)

in2 <- read_csv("imageJ_measurements.csv")
in2$strain2 <- str_replace_all(in2$strain, c("trmBtbsP" = "mut", "tbsP" = "mut", "trmB" = "mut"))
in2
```

```{r, message=FALSE, warning=FALSE}
in2 %>%
  group_by(filename, plate, condition, strain2) %>%
  summarise(average = mean(area_in)) %>%
  filter(condition != "tbsP+noglu" | plate != 3) %>%
  group_by(filename, plate, condition) %>%
  summarise(ratio = average[strain2 == "mut"] / average[strain2 == "WT"]) %>%
  group_by(condition) %>%
  summarise(
    avg_ratio = mean(ratio),
    sd_ratio = sd(ratio)
  ) %>%
  separate(condition, c("strain", "condition")) -> tmp2

ggplot(data = tmp2, aes(y = avg_ratio, x = condition, fill = strain)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = avg_ratio - sd_ratio, ymax = avg_ratio + sd_ratio),
    width = .1, size = 1, position = position_dodge(.9)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02))) +
  theme_bw()

ggplot(data = tmp2, aes(y = avg_ratio, x = strain, fill = condition)) +
  geom_hline(yintercept = 1, color = "grey", linetype = "dashed", size = 1) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = avg_ratio - sd_ratio, ymax = avg_ratio + sd_ratio),
    width = .1, size = 1, position = position_dodge(.9)
  ) +
  ylab("proportion WT diameter") +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.02)),
    limits = c(0, 6.5)
  ) +
  theme_bw() +
  theme(legend.position = c(.9, .85))
```
there are relatively few differences between the strains and WT when glucose is added to the media. Even trmB-glu cells are motile.largest motility difference in tbsP-glu, suggesting that tbsP represses motility in -glu conditions. Double deletion phenocopies. 

look at outliers
```{r, fig.width=12, warning=FALSE, message=FALSE}
in2 %>%
  group_by(filename, plate, condition, strain2) %>%
  summarise(average = mean(area_in)) %>%
  filter(condition != "tbsP+noglu" | plate != 3) %>%
  separate(condition, c("strain", "condition")) -> tmp3

tmp3 %>%
  ggplot(., ) +
  geom_col_pattern(aes(y = average, x = condition, fill = as.character(plate), pattern_fill = strain2),
    stat = "identity", position = "dodge",
    colour  = "black"
  ) +
  facet_wrap(~strain) +
  theme_bw()
```
tbsP+glu plate 1
trmB+glu and noglu plate 3
trmBtbsP+glu plate 1
tbsP+noglu plate 1

```{r, message=FALSE, warning=FALSE}
in2 %>%
  group_by(filename, plate, condition, strain2) %>%
  summarise(average = mean(area_in)) %>%
  filter(condition != "tbsP+noglu" | plate != 3) %>%
  group_by(filename, plate, condition) %>%
  summarise(ratio = average[strain2 == "mut"] / average[strain2 == "WT"]) -> rats

t.test(filter(rats, condition == "tbsP+glu")$ratio, filter(rats, condition == "tbsP+noglu")$ratio)
t.test(filter(rats, condition == "trmBtbsP+glu")$ratio, filter(rats, condition == "trmBtbsP+noglu")$ratio)
t.test(filter(rats, condition == "trmB+glu")$ratio, filter(rats, condition == "trmB+noglu")$ratio)
```

tbsP and trmBtbsP are sig different between plus and minus glucose

# Make final figure for the manuscript
```{r, message=FALSE, warning=FALSE}
strain.colors <- c(trmB = "#35B779", trmBtbsP = "#A6A560", tbsP = "#E79A51")
condition.colors <- c(glu = "grey80", noglu = "white")

in2 %>%
  group_by(filename, plate, condition, strain2) %>%
  summarise(average = mean(area_in)) %>%
  filter(condition != "tbsP+noglu" | plate != 3) %>%
  group_by(filename, plate, condition) %>%
  summarise(ratio = average[strain2 == "mut"] / average[strain2 == "WT"]) %>%
  group_by(condition) %>%
  summarise(
    avg_ratio = mean(ratio),
    sd_ratio = sd(ratio)
  ) %>%
  separate(condition, c("strain", "condition")) -> tmp2

pdf("motility.graph.pdf", width = 4, height = 4)

ggplot(data = tmp2, aes(y = avg_ratio, x = strain, fill = condition, color = strain)) +
  geom_hline(yintercept = 1, color = "grey", linetype = "dashed", size = 1) +
  geom_bar(stat = "identity", position = "dodge", size = 1.5) +
  geom_errorbar(aes(ymin = avg_ratio - sd_ratio, ymax = avg_ratio + sd_ratio),
    width = .1, size = 1, position = position_dodge(.9)
  ) +
  ylab("proportion WT diameter") +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.02)),
    limits = c(0, 6.5)
  ) +
  theme_bw() +
  theme(legend.position = c(.9, .85)) +
  scale_color_manual(values = strain.colors) +
  scale_fill_manual(values = condition.colors)

dev.off()
```
